Bayesian determination of the effect of a deep eutectic solvent on the structure of lipid monolayers

### 2018-10-01 reports/preprint.tex
After tom's clean up job

\begin{abstract}
  Deep eutectic solvents present a novel class of non-aqueous room temperature solvent with tunable properties, that are capable of promoting the self-assembly of surfactant molecules.
  However, the solvation model in these systems still challenges the classic understanding of amphiphilicity.
  In this work, we present the first example of the self-assembly of phospholipid monolayers at the interface between air and a non-aqueous solvent.
  Furthermore, we use novel, chemically-consistent Bayesian modelling of X-ray and neutron reflectometry measurements to show the ability of the deep eutectic solvent to interact with the phosphatidylglycerol lipid head, leading to an apparent increase in the volume compared to that observed in water.
  No such change was observed for the phosphocholine head, indicating that the interaction is head specific.
\begin{description}
\item[Usage]
Electronic Supplementary Information (ESI) available: All analysis/plotting scripts and figure files, allowing for a fully reproducible, and automated, analysis workflow for the work presented is available at \url{https://github.com/arm61/lipids_at_airdes} (DOI: **some doi**) under an MIT license. Reduced experimental datasets are available at \url{https://researchdata.bath.ac.uk/id/eprint/548}, under a CC-BY 4.0 license.
\end{description}
\end{abstract}

Deep eutectic solvents (DES) are a green, sustainable solvent system obtained through the complexation of naturally occurring compounds, such as sugars, alcohols, amines, and carboxylic acids\cite{Smith2014,Dai2013}.
An extensive hydrogen bonding network is present, allowing the mixture to remain liquid at room temperature\cite{Hammond2016,Hammond2017,Araujo2017}.
Through different combinations of the precursor materials, it is possible to tune the solvent's physicochemical properties, such as polarity\cite{Pandey2014}, viscosity and surface tension\cite{Smith2014}, network charge\cite{Zahn2016}, and hydrophobicity\cite{Ribeiro2015,vanOsch2015}.
Furthermore, recently DES have indicated the presence of a solvophobic effect through the promotion of surfactant micelle formation\cite{Sanchez-Fernandez2016,Arnold2015}, phospholipid bilayer formation\cite{Bryant2017,Bryant2016,Gutierrez2009}, and the ability to stabilise non-ionic polymer conformations\cite{Sapir2016} and protein conformation\cite{Sanchez-Fernandez2017}.

Phospholipid monolayer formation at the air/liquid interface plays a key role in many biological and technological processes, where the solvent-specific solubility of different components of the phospholipid enables a stable monolayer to form at the interface\cite{Mohwald1990}.
Investigations at the air/salt-water interface have identified the importance that interactions between charged phospholipid heads and ions present in solution have on the structure, monomer packing and stability of the monolayer\cite{Mohwald1990,Kewalramani2010}.
Despite the broad interest in these systems, the presence of stable phospholipid monolayers at the interface between air and a non-aqueous media has not been previously reported, to the best of the authors' knowledge.

This work presents the first investigation of the structure of four phospholipid monolayers at the air-DES interface, as determined by chemically-consistent modelling of X-ray reflectometry (XRR) measurements.
Evaluating the interaction of two, chemically distinct, phospholipid heads with this non-aqueous solvent, in addition to the effect of the tail chain length. The analysis was then extended to model complementary neutron reflectometry (NR) measurements for two contrasts of DMPC and DPPC at a single surface pressure.

Recent developments in computational resources and software have enabled powerful methodologies and algorithms to be harnessed by those from non-expert backgrounds.
This has benefitted significantly from open-source software projects such as the Python language\cite{vanRossum1995} and the Jupyter notebooks framework\cite{Kluyver2016}.
In the area of NR and XRR data-analysis, the landscape of open-source software is diverse, with a range of software packages available from a variety of sources; refnx\cite{Nelson2018}, MOTOFIT\cite{Nelson2006}, Aurore\cite{Gerelli2016}, and GenX\cite{Bjorck2007}.

@@

%
\begin{figure}
\centering
\includegraphics[width=0.3\textwidth]{figures/head_groups}
\caption{\label{fig:heads}\small The two lipid classes with different head groups compared in this study, where R indicates the hydrocarbon tail; (a) phosphatidylglycerol (PG), (b) phosphocholine (PC).}
\end{figure}
%
We developed a chemically-consistent model (detail in the ESI) that allowed for the co-refinement of XRR and NR measurements at different surface concentration.
The model was applied to the study of four phospholipids monolayers, namely 1,2-dipalmitoyl-sn-glycero-3-phosphocholine (DPPC, C$_{16}$ tails), 1,2-dimyristoyl-sn-glycero-3-phosphocholine (DMPC, C$_{14}$ tails),  1,2-dilauroyl-sn-glycero-3-phosphocholine (DLPC, C$_{12}$ tails) and 1,2-dimyristoyl-sn-glycero-3-phospho-(1'-rac-glycerol) (DMPG, C$_{14}$ tails), at the air-1:2 choline choride:glycerol mixture interface, by XRR at a series of different surface pressures, between \SIrange{15}{40}{\milli\newton\per\meter}.
Unlike monolayer models applied previously\cite{Mohwald1990,Kewalramani2010,Bayerl1990,Johnson1991,Clifton2012,Helm1987,Daillant1990}, our model made no assumption of the volume of the lipid head, $V_h$, or tail, $V_t$.
Instead these parameters, along with the thickness of the phospholipid head layer, $d_h$, were allowed to vary for each lipid.
However, their values were constained to be self-consistent for a single lipid over multiple measurements at different surface pressures.

This model was required, despite the general consensus that the volume of the phosphocholine (PC) head is \SIrange{320}{360}{\cubic\angstrom}, while the phosphatidylglycerol (PG) head is \SIrange{289}{291}{\cubic\angstrom} (see Table \ref{tab:water}), as it not clear if the presence of the DES would affect the head volume due to electrostatic interactions, which may differ from those between the head and water.
Furthermore, the effect of pressure, which is not necessarily accounted for the in the literature, has been shown to compress the lipid tail volume\cite{Marsh2010,Small1984}, and the need to account for this hydrocarbon compaction according to their phase in the modelling of NR data has recently been demonstrated\cite{Campbell2018}.
%
\begin{figure}
	\centering
	\includegraphics[width=0.45\textwidth]{figures/dlpc_ref_sld}
	\includegraphics[width=0.45\textwidth]{figures/dmpc_ref_sld}
	\includegraphics[width=0.45\textwidth]{figures/dppc_ref_sld}
	\includegraphics[width=0.45\textwidth]{figures/dmpg_ref_sld}
	\caption{\small The XRR profiles (left) and SLD profiles (right) for each of the four lipids; (a) DLPC, (b) DMPC, (c) DPPC, (d) DMPG, at the four measured surface pressures; see legend above each plot. The different surface pressure XRR profiles have been offset in the $y$-axis by an order of magnitude and SLD profiles offset in the $y$-axis by \SI{5e-6}{\per\square\angstrom}, for clarity.}
	\label{fig:lipids}
\end{figure}
%

The custom model for each lipid, fitted to the experimental XRR data, and the associated SLD profiles are shown in Figure \ref{fig:lipids}.
Table \ref{tab:liptab} presents the mean value, and asymmetric uncertainies that correspond to a \SI{95}{\percent} confidence interval, for the PDF for each of the varying parameters; the tail tile angle, $\theta_t$, and the interfacial roughness, $\sigma$, which were free to vary independent of surface pressure.
In addition to the head and tail volumes, $V_h$ and $V_t$ respectively, and the head layer thickness, $d_h$, for which a single value was fitted for each lipid across all surface pressures.
%
\begin{table*}
  \centering
	\caption{\label{tab:liptab}\small  The best-fit values, and associated 95 \% confidence intervals for the varying parameters in the XRR models, at the highest surface pressure (SP) measured. The values of $d_t$ were found from the appropriate values of $\theta_t$ using Eqn. \ref{equ:tl} and the values for $\phi_h$ were obtained from the appropriate use of Eqn. \ref{equ:phih}.}
	\begin{tabular}{ccccc}
		Lipid & DLPC & DMPC & DPPC & DMPG \\
    SP/\si{\milli\newton\per\meter} & 35 & 40 & 30 & 30 \\
		\hline
		$\theta_t$/\si{\degree} & \input{../output/dlpc/angle35.txt} & \input{../output/dmpc/angle40.txt} & \input{../output/dppc/angle30.txt} & \input{../output/dmpg/angle30.txt} \\
		$\sigma$/\si{\angstrom} & \input{../output/dlpc/rough35.txt} & \input{../output/dmpc/rough40.txt} & \input{../output/dppc/rough30.txt} & \input{../output/dmpg/rough30.txt} \\
    \hline
    $V_t$/\si{\cubic\angstrom} & \input{../output/dlpc/vt.txt} & \input{../output/dmpc/vt.txt} & \input{../output/dppc/vt.txt} & \input{../output/dmpg/vt.txt} \\
		$V_h$/\si{\cubic\angstrom} & \input{../output/dlpc/vh.txt} & \input{../output/dmpc/vh.txt} & \input{../output/dppc/vh.txt} & \input{../output/dmpg/vh.txt} \\
		$d_h$/\si{\angstrom} & \input{../output/dlpc/head.txt} & \input{../output/dmpc/head.txt} & \input{../output/dppc/head.txt} & \input{../output/dmpg/head.txt} \\
    \hline
    $\phi_h$/$\times10^{-2}$ & \input{../output/dlpc/solh35.txt} & \input{../output/dmpc/solh40.txt} & \input{../output/dppc/solh30.txt} & \input{../output/dmpg/solh30.txt} \\
		$d_t$/\si{\angstrom} & \input{../output/dlpc/tail35.txt} & \input{../output/dmpc/tail40.txt} & \input{../output/dppc/tail30.txt} & \input{../output/dmpg/tail30.txt} \\
	\end{tabular}
\end{table*}
%

In keeping this previous literature, \cite{Mohwald1990,Vaknin1991}, the tail layer thickness, $d_t$, increases with the number of hydrocarbon atoms present.
Furthermore, the thickness of the tail layers agress well with values found for water-analogues; \input{../output/dmpc/tail30.txt}\si{\angstrom} at \SI{30}{\milli\newton\per\meter} in DES compared with $d_t=\SI{15.8}{\angstrom}$ at \SI{30}{\milli\newton\per\meter}\cite{Johnson1991} in water for DMPC, and \input{../output/dppc/tail30.txt}\si{\angstrom} at \SI{30}{\milli\newton\per\meter} in DES compared with $d_t=\SI{16.7}{\angstrom}$ at \SI{40}{\milli\newton\per\meter}\cite{Helm1987} in water for DPPC.

Figure \ref{fig:lipresults} shows the tail layer thickness increasing with surface pressure, before plateuing; for DPPC this occurs at \SI{20}{\milli\newton\per\meter}, DMPC at \SI{30}{\milli\newton\per\meter} and for DMPG and DLPC the plateu can be assumed to be at higher pressures than those studied.
This phenomenon has been noted before for DMPC\cite{Bayerl1990} and DPPC\cite{Campbell2018} at the air-water interface.
%
\begin{figure}
	\centering
	\includegraphics[width=0.45\textwidth]{figures/dlpc_vh_dt_phi}
	\includegraphics[width=0.45\textwidth]{figures/dmpc_vh_dt_phi}
	\includegraphics[width=0.45\textwidth]{figures/dppc_vh_dt_phi}
	\includegraphics[width=0.45\textwidth]{figures/dmpg_vh_dt_phi}
	\caption{\small The PDFs of the head volume (left) and variation of $d_t$ (squares) and $\phi_h$ (circles) with surface pressure for each of the four lipids; (a) DLPC, (b) DMPC, (c) DPPC, (d) DMPG. The values of $d_t$ were found from the appropriate values of $\theta_t$ using Eqn. \ref{equ:tl}.}
	\label{fig:lipresults}
\end{figure}
%

@@

Due to the poor wetting properties of the DES, it was not possible to collect surface isotherm measurements.
However, grazing incidence X-ray diffraction was collected for DMPC and DPPC at \SI{30}{\milli\newton\per\meter}, indicating that the lipids had adopted the liquid condensed phase (see Section \ref{sec:gixd}).
This is the same phase behaviour as is observed in water \cite{Phillips1968} and by considering that there will be little interaction between the lipid tails and the DES it is unlikely that a significant change in the phase behaviour \cite{Nagle1976}.
The means that at all four surface pressures, as is the case on water, the lipids would adopt the same liquid condensed phase and any variation in the structure with surface pressure would manifest only as a change in the tail thickness, via the tail tilt angle.
This allowed a single tail volume to be fitted for each lipid at all four surface pressures that were measured.
It is clear when comparing Tables \ref{tab:water} and \ref{tab:liptab} that some of the tail volumes are reduced in the current XRR measurements compared to those determined previously.
The reduction was found to be \SIrange{8}{12}{\percent} for DPPC, DMPC and DLPC when compared with literature sources at \SIrange{24}{30}{\celsius}, this is in good agreement with the maximum compression percentage of \SI{15}{\percent} noted by Small \emph{et al.}\cite{Small1984}. DMPG shows a small increase in the tail volume relative to the literature value quoted at lower temperature.
Notably our value is similar to that found for DMPC on DES, which has the same tail structure and suggests that our results are at least self-consistent.

@@

The determined volumes for the head and tails, and the thickness of the head layers, for DMPC and DPPC were then used as constraints in the co-refinement of the custom model against two contrasts NR data (Figure \ref{fig:neutron}). In these analysis the only two varying parameters were the tail tilt angle, $\theta_t$, and the interfacial roughness, $\sigma$.
The quality of agreement with the NR measurements indicates that the values found for the head and tail volumes are consistent between the pair of measurements for the same system.
It is clear, that again stable monolayers of the lipids are forming at the air-DES interface, and that the head and tail volumes determined from XRR measurements are robust-enough to be used in the modelling of NR data.
Futhermore, the trends present with varying surface pressure in the XRR models are consistent with that found in the NR models.
%
\begin{figure}
	\centering
	\includegraphics[width=0.45\textwidth]{figures/dmpc_20n_ref_sld}
	\includegraphics[width=0.45\textwidth]{figures/dppc_20n_ref_sld}
	\caption{\small The NR and SLD profiles at a surface pressure of 20 mNm$^{-1}$ for two contrasts (see legend above each plot); (a) DMPC, (b) DPPC. The NR profiles have been offset in the $y$-axis by an order of magnitude and SLD profiles offset in the $y$-axis by \SI{5e-6}{\per\square\angstrom}, for clarity.}
	\label{fig:neutron}
\end{figure}
%
%
\begin{table*}
  \centering
	\caption{\label{tab:neutron}\small The best-fit values, and associated 95 \% confidence intervals for the varying parameters in the co-refined NR models. The values of $d_t$ were found from the appropriate values of $\theta_t$ using Eqn. \ref{equ:tl}, and the values of $\phi_h$ were found using Eqn. \ref{equ:phih}.}
	\begin{tabular}{ccccc}
    Lipid & \multicolumn{2}{c}{d$_{54}$-DMPC} & \multicolumn{2}{c}{d$_{62}$-DPPC} \\
    SP/mNm$^{-1}$ & 20 & 25 & 15 & 20 \\
		\hline
		$\theta_t$/$^\circ$ & \input{../output/dmpc/angle20_neutron.txt} & \input{../output/dmpc/angle25_neutron.txt} & \input{../output/dppc/angle15_neutron.txt} & \input{../output/dppc/angle20_neutron.txt} \\
		$\sigma_{t,h,s}$/\AA & \input{../output/dmpc/rough20_neutron.txt} & \input{../output/dmpc/rough25_neutron.txt} & \input{../output/dppc/rough15_neutron.txt} & \input{../output/dppc/rough20_neutron.txt} \\
    \hline
    $\phi_h$/$\times10^{-2}$ & \input{../output/dmpc/solh20_neutron.txt} & \input{../output/dmpc/solh25_neutron.txt} & \input{../output/dppc/solh15_neutron.txt} & \input{../output/dppc/solh20_neutron.txt} \\
		$d_t$/\AA & \input{../output/dmpc/tail20_neutron.txt} & \input{../output/dmpc/tail25_neutron.txt} & \input{../output/dppc/tail15_neutron.txt} & \input{../output/dppc/tail20_neutron.txt} \\
	\end{tabular}
\end{table*}
%

@@

A physically and chemically constrained modelling approach and Bayesian analysis method was used to rationalise these measurements showing that the structures are remarkably similar at the air-DES interface to those previously observed at the air-water interface.
This has the important implication that DES therefore offers the possibility of performing studies of model membranes in the absence of water.
Such applications may include fundamental investigations of phospholipid monolayers in extreme environments (total or partial absence of water, cryogenic temperatures), protein membrane interactions and development of new technologies for drug delivery.
However, the fact remains that the PG head containing lipid did show a significant difference; having a larger head volume than observed for the same system in water.
This shows that the transfer of lipids to a DES is not just a simple substitution of the subphase. In this specific case we have proposed an explanation based on unfolding of the PG head that is enabled by electrostatic screening of the head charges by the charged solvent.

@@

\setlength{\bibsep}{1pt}
\bibliography{rsc}% Produces the bibliography via BibTeX.
\bibliographystyle{rsc}
\end{document}

>>>>

\begin{abstract}
Deep eutectic solvents present a novel class of non-aqueous room temperature solvent with tunable properties, that are capable of promoting the self-assembly of surfactant molecules.
However, the solvation model in these systems still challenges the classic understanding of amphiphilicity.
In this work, we present the first example of the self-assembly of phospholipid monolayers at the interface between air and a non-aqueous solvent.
We use novel, chemically-consistent Bayesian modelling of X-ray and neutron reflectometry measurements to show that these monolayers broadly behave as they do on water.
However, the ability of the deep eutectic solvent to interact with the phosphatidylglycerol lipid head, leads to an apparent increase in its volume compared to that observed in water.
No such change was observed for the phosphocholine head, indicating that such interactions are head, and therefore solvent, specific.
\begin{description}
\item[Usage]
Electronic Supplementary Information (ESI) available: All analysis/plotting scripts and figure files, allowing for a fully reproducible, and automated, analysis workflow for the work presented is available at \url{https://github.com/arm61/lipids_at_airdes} (DOI: **some doi**) under an MIT license. Reduced experimental datasets are available at \url{https://researchdata.bath.ac.uk/id/eprint/548}, under a CC-BY 4.0 license.
\end{description}
\end{abstract}

Deep eutectic solvents (DES) are green, sustainable liquids that are obtained through the combination of ionic species with compounds that act as hydrogen bond donors, such as sugars, alcohols, amines, and carboxylic acids\cite{Smith2014,Dai2013}.
The resulting extensive hydrogen bonding network is able to stabilise the ionic species and allows the eutectic mixture to remain liquid at room temperature\cite{Hammond2016,Hammond2017,Araujo2017}.
Through different combinations of the precursor materials, it is possible to tune the solvent's physicochemical properties, such as polarity\cite{Pandey2014}, viscosity and surface tension\cite{Smith2014}, network charge\cite{Zahn2016}, and hydrophobicity\cite{Ribeiro2015,vanOsch2015}.
Recently DES have also been shown to exhibit a "solvophobic" effect through the promotion of surfactant micelle formation\cite{Sanchez-Fernandez2016,Arnold2015}, phospholipid bilayer formation\cite{Bryant2017,Bryant2016,Gutierrez2009}, and the ability to stabilise non-ionic polymer\cite{Sapir2016} and protein conformations\cite{Sanchez-Fernandez2017}.

Phospholipid monolayers at the air/water interface have been widely studied as simplistic models for biological membranes.
As such, they have been used to gain insight into many biological processes that are technologically and medically relevant.
For example, investigations at the air/salt-water interface have identified the importance that interactions between charged phospholipid heads and ions present in solution have on the structure, monomer packing and stability of the monolayer\cite{Mohwald1990,Kewalramani2010}.
However, the native environment for lipids in-vivo is far from a simple aqueous solution.
In fact, it has been suggested\cite{Choi2011,Hammond2017} that DES effectively form within the crowded cellular environment and could assist in solubilizing biological species in an intermediate environment between that of the hydrophobic phospholipid tails and highly polar water rich regions, particularly assisting survival under extreme conditions such as freezing temperatures or drought where the water content of cells is restricted.
Despite the broad interest in these systems, the presence of stable phospholipid monolayers at the interface between air and any non-aqueous media has not been previously reported, to the best of the authors' knowledge.

This work presents the first investigation of the structure of phospholipid monolayers at an air-DES interface.
We have used a chemically-consistent approach to model X-ray (XRR) and neutron (NR) reflectometry measurements and thereby evaluate the effect of this non-aqueous solvent on the structure of phospholipid monolayers.

Recent developments in computational resources and software have enabled powerful methodologies and algorithms to be harnessed by those from non-expert backgrounds.
This has benefitted significantly from open-source software projects such as the Python language\cite{vanRossum1995} and the Jupyter notebooks framework\cite{Kluyver2016}.
In the area of NR and XRR, the landscape of data-analysis software is diverse, with a range of software packages available from a variety of sources; refnx\cite{Nelson2018}, MOTOFIT\cite{Nelson2006}, Rascal\cite{HughesRascal} Aurore\cite{Gerelli2016}, and GenX\cite{Bjorck2007}.

@@

%
\begin{figure}
\centering
\includegraphics[width=0.3\textwidth]{figures/head_groups}
\caption{\label{fig:heads}\small The two lipid classes with different head groups compared in this study, where R indicates the hydrocarbon tail; (a) phosphatidylglycerol (PG), (b) phosphocholine (PC).}
\end{figure}
%
We developed a chemically-consistent model (detail in the ESI) that allowed for the co-refinement of reflectometry measurements at different surface concentration.
The model was applied to the study of four phospholipids monolayers, namely 1,2-dipalmitoyl-sn-glycero-3-phosphocholine (DPPC, C$_{16}$ tails), 1,2-dimyristoyl-sn-glycero-3-phosphocholine (DMPC, C$_{14}$ tails),  1,2-dilauroyl-sn-glycero-3-phosphocholine (DLPC, C$_{12}$ tails) and 1,2-dimyristoyl-sn-glycero-3-phospho-(1'-rac-glycerol) (DMPG, C$_{14}$ tails), at the air-1:2 choline choride:glycerol mixture interface, by XRR at a series of different surface pressures, between \SIrange{15}{40}{\milli\newton\per\meter}.
Unlike monolayer models applied previously\cite{Mohwald1990,Kewalramani2010,Bayerl1990,Johnson1991,Clifton2012,Helm1987,Daillant1990}, our model made no assumption of the volume of the lipid head, $V_h$, or tail, $V_t$.
Instead these parameters, along with the thickness of the phospholipid head layer, $d_h$, were allowed to vary for each lipid.
However, their values were constained to be self-consistent for a single lipid over multiple measurements at different surface pressures.

This model was required, despite the general consensus that the volume of the phosphocholine (PC) head is \SIrange{320}{360}{\cubic\angstrom}, while the phosphatidylglycerol (PG) head is \SIrange{289}{291}{\cubic\angstrom} (see Table \ref{tab:water}), since the DES may affect the head volume, because of differing electrostatic interactions compared to water.
Furthermore, it is known that, on water, increased surface pressure and associated liquid-expanded to liquid-condensed (LE-LC) phase transitions lead to a compression of the lipid tail volume\cite{Marsh2010,Small1984}.
It has been shown recently that such hydrocarbon compaction should be included when modelling reflectometry data\cite{Campbell2018}, something that has not necessarily been accounted for the in the literature.

%
\begin{figure}
	\centering
	\includegraphics[width=0.45\textwidth]{figures/dlpc_ref_sld}
	\includegraphics[width=0.45\textwidth]{figures/dmpc_ref_sld}
	\includegraphics[width=0.45\textwidth]{figures/dppc_ref_sld}
	\includegraphics[width=0.45\textwidth]{figures/dmpg_ref_sld}
	\caption{\small The XRR profiles (left) and SLD profiles (right) for each of the four lipids; (a) DLPC, (b) DMPC, (c) DPPC, (d) DMPG, at the four measured surface pressures; see legend above each plot. The different surface pressure XRR profiles have been offset in the $y$-axis by an order of magnitude and SLD profiles offset in the $y$-axis by \SI{5e-6}{\per\square\angstrom}, for clarity.}
	\label{fig:lipids}
\end{figure}
%

The custom model for each lipid, based on the standard two-layer model widely used for lipids on water, was first fitted to the experimental XRR data, and the associated SLD profiles are shown in Figure \ref{fig:lipids}.
Table \ref{tab:liptab} presents the mean value, and asymmetric uncertainies that correspond to a \SI{95}{\percent} confidence interval, for the PDF for each of the varying parameters; the tail tile angle, $\theta_t$, the interfacial roughness, $\sigma$, the head and tail volumes, $V_h$ and $V_t$ respectively, and the head layer thickness, $d_h$.
In order to constrain the fitting process only $\theta_t$ and $\sigma$ were allowed to vary independently of surface pressure. The other parameters, $V_h$, $V_t$ and $d_h$ were fitted to a single value for all of the surface pressures measured for each lipid.

%
\begin{table*}
  \centering
	\caption{\label{tab:liptab}\small The best-fit values, and associated 95 \% confidence intervals for the varying parameters in the XRR models, at the highest surface pressure (SP) measured. The values of $d_t$ were found from the appropriate values of $\theta_t$ using Eqn. \ref{equ:tl} and the values for $\phi_h$ were obtained from the appropriate use of Eqn. \ref{equ:phih}.}
	\begin{tabular}{ccccc}
		Lipid & DLPC & DMPC & DPPC & DMPG \\
    SP/\si{\milli\newton\per\meter} & 35 & 40 & 30 & 30 \\
		\hline
		$\theta_t$/\si{\degree} & \input{../output/dlpc/angle35.txt} & \input{../output/dmpc/angle40.txt} & \input{../output/dppc/angle30.txt} & \input{../output/dmpg/angle30.txt} \\
		$\sigma$/\si{\angstrom} & \input{../output/dlpc/rough35.txt} & \input{../output/dmpc/rough40.txt} & \input{../output/dppc/rough30.txt} & \input{../output/dmpg/rough30.txt} \\
    \hline
    $V_t$/\si{\cubic\angstrom} & \input{../output/dlpc/vt.txt} & \input{../output/dmpc/vt.txt} & \input{../output/dppc/vt.txt} & \input{../output/dmpg/vt.txt} \\
		$V_h$/\si{\cubic\angstrom} & \input{../output/dlpc/vh.txt} & \input{../output/dmpc/vh.txt} & \input{../output/dppc/vh.txt} & \input{../output/dmpg/vh.txt} \\
		$d_h$/\si{\angstrom} & \input{../output/dlpc/head.txt} & \input{../output/dmpc/head.txt} & \input{../output/dppc/head.txt} & \input{../output/dmpg/head.txt} \\
    \hline
    $\phi_h$/$\times10^{-2}$ & \input{../output/dlpc/solh35.txt} & \input{../output/dmpc/solh40.txt} & \input{../output/dppc/solh30.txt} & \input{../output/dmpg/solh30.txt} \\
		$d_t$/\si{\angstrom} & \input{../output/dlpc/tail35.txt} & \input{../output/dmpc/tail40.txt} & \input{../output/dppc/tail30.txt} & \input{../output/dmpg/tail30.txt} \\
	\end{tabular}
\end{table*}
%

In keeping this previous literature, \cite{Mohwald1990,Vaknin1991}, the tail layer thickness, $d_t$, increases with the number of hydrocarbon atoms present.
Furthermore, the thickness of the tail layers agrees well with values found for water-analogues; \input{../output/dmpc/tail30.txt}\si{\angstrom} at \SI{30}{\milli\newton\per\meter} in DES compared with $d_t=\SI{15.8}{\angstrom}$ at \SI{30}{\milli\newton\per\meter}\cite{Johnson1991} in water for DMPC, and \input{../output/dppc/tail30.txt}\si{\angstrom} at \SI{30}{\milli\newton\per\meter} in DES compared with $d_t=\SI{16.7}{\angstrom}$ at \SI{40}{\milli\newton\per\meter}\cite{Helm1987} in water for DPPC.
In all cases the surface roughness was found to be relatively high when compared to water. However this should be considered in the context that the absolute surface tension of the DES is less than water with a corresponding higher surface roughness\cite{Arnold2015}.

Figure \ref{fig:lipresults} shows the tail layer thickness increasing with surface pressure, before plateuing; for DPPC this occurs at \SI{20}{\milli\newton\per\meter}, DMPC at \SI{30}{\milli\newton\per\meter} and for DMPG and DLPC the plateu can be assumed to be at higher pressures than those studied.
This phenomenon has been noted before for DMPC\cite{Bayerl1990} and DPPC\cite{Campbell2018} at the air-water interface.

%
\begin{figure}
	\centering
	\includegraphics[width=0.45\textwidth]{figures/dlpc_vh_dt_phi}
	\includegraphics[width=0.45\textwidth]{figures/dmpc_vh_dt_phi}
	\includegraphics[width=0.45\textwidth]{figures/dppc_vh_dt_phi}
	\includegraphics[width=0.45\textwidth]{figures/dmpg_vh_dt_phi}
	\caption{\small The PDFs of the head volume (left) and variation of $d_t$ (squares) and $\phi_h$ (circles) with surface pressure for each of the four lipids; (a) DLPC, (b) DMPC, (c) DPPC, (d) DMPG. The values of $d_t$ were found from the appropriate values of $\theta_t$ using Eqn. \ref{equ:tl}.}
	\label{fig:lipresults}
\end{figure}
%

@@

While we have confidence that the individual surface pressures measured were reliable, we were unable to collect consistent Langmuir isotherm measurements, due to the poor wetting properties of the DES.
This inconsistency means that we cannot be confident in the phase of the lipids based on the surface pressure alone.
Instead we have used grazing incidence X-ray diffraction to confirm the phases of DMPC and DPPC at \SI{30}{\milli\newton\per\meter}.
Thus DPPC was found to be in the LC phase at room temperature while DMPC was in the LE phase at room temperature and the LC phase at 7C (see Section \ref{sec:gixd}).
This is essentially the same phase behaviour as is observed in water \cite{Phillips1968} and by considering that there is little interaction between the lipid tails and the DES we believe that it is unlikely that there is a significant change in the phase behaviour in these systems\cite{Nagle1976}.
We therefore assume that for all four surface pressures, as is the case on water, the lipids adopt the same phase for all surface pressures measured and any variation in the structure would manifest only as a change in the tail thickness, via the tail tilt angle.
This allowed a single tail volume to be fitted for each lipid at all four surface pressures that were measured.

It is clear when comparing Tables \ref{tab:water} and \ref{tab:liptab} that some of the tail volumes are reduced in the current XRR measurements compared to those determined previously.
The reduction was found to be \SIrange{8}{12}{\percent} for DPPC, DMPC and DLPC when compared with literature sources at \SIrange{24}{30}{\celsius}, this is in good agreement with the maximum compression percentage of \SI{15}{\percent} noted by Small \emph{et al.}\cite{Small1984}. DMPG shows a small increase in the tail volume relative to the literature value quoted at lower temperature.
Notably our value is similar to that found for DMPC on DES, which has the same tail structure and suggests that our results are at least self-consistent.

@@

The determined volumes for the head and tails, and the thickness of the head layers, for DMPC and DPPC were then used as constraints in the co-refinement of the custom model against two contrasts NR data (Figure \ref{fig:neutron}). In these analysis the only two varying parameters were the tail tilt angle, $\theta_t$, and the interfacial roughness, $\sigma$.
The quality of agreement with the NR measurements indicates that the values found for the head and tail volumes are consistent between the pair of measurements for the same system.
It is clear, that again stable monolayers of the lipids are forming at the air-DES interface, and that the head and tail volumes determined from XRR measurements are robust-enough to be used in the modelling of NR data.
Futhermore, the trends present with varying surface pressure in the XRR models are consistent with that found in the NR models.

%
\begin{figure}
	\centering
	\includegraphics[width=0.45\textwidth]{figures/dmpc_20n_ref_sld}
	\includegraphics[width=0.45\textwidth]{figures/dppc_20n_ref_sld}
	\caption{\small The NR and SLD profiles at a surface pressure of 20 mNm$^{-1}$ for two contrasts (see legend above each plot); (a) DMPC, (b) DPPC. The NR profiles have been offset in the $y$-axis by an order of magnitude and SLD profiles offset in the $y$-axis by \SI{5e-6}{\per\square\angstrom}, for clarity.}
	\label{fig:neutron}
\end{figure}
%
%
\begin{table*}
  \centering
	\caption{\label{tab:neutron}\small The best-fit values, and associated 95 \% confidence intervals for the varying parameters in the co-refined NR models. The values of $d_t$ were found from the appropriate values of $\theta_t$ using Eqn. \ref{equ:tl}, and the values of $\phi_h$ were found using Eqn. \ref{equ:phih}.}
	\begin{tabular}{ccccc}
    Lipid & \multicolumn{2}{c}{d$_{54}$-DMPC} & \multicolumn{2}{c}{d$_{62}$-DPPC} \\
    SP/mNm$^{-1}$ & 20 & 25 & 15 & 20 \\
		\hline
		$\theta_t$/$^\circ$ & \input{../output/dmpc/angle20_neutron.txt} & \input{../output/dmpc/angle25_neutron.txt} & \input{../output/dppc/angle15_neutron.txt} & \input{../output/dppc/angle20_neutron.txt} \\
		$\sigma_{t,h,s}$/\AA & \input{../output/dmpc/rough20_neutron.txt} & \input{../output/dmpc/rough25_neutron.txt} & \input{../output/dppc/rough15_neutron.txt} & \input{../output/dppc/rough20_neutron.txt} \\
    \hline
    $\phi_h$/$\times10^{-2}$ & \input{../output/dmpc/solh20_neutron.txt} & \input{../output/dmpc/solh25_neutron.txt} & \input{../output/dppc/solh15_neutron.txt} & \input{../output/dppc/solh20_neutron.txt} \\
		$d_t$/\AA & \input{../output/dmpc/tail20_neutron.txt} & \input{../output/dmpc/tail25_neutron.txt} & \input{../output/dppc/tail15_neutron.txt} & \input{../output/dppc/tail20_neutron.txt} \\
	\end{tabular}
\end{table*}
%

@@

A physically and chemically constrained modelling approach and Bayesian analysis method was used to rationalise these measurements showing that the structures are remarkably similar at the air-DES interface to those previously observed at the air-water interface.
This has the important implication that DES therefore offers the possibility of performing studies of model membranes in the absence of water.
Such applications may include fundamental investigations of phospholipid monolayers in extreme environments (total or partial absence of water, cryogenic temperatures), protein membrane interactions and development of new technologies for drug delivery.
However, the fact remains that the PG lipid did show a significant difference; having a larger head volume than observed for the same system in water.
This shows that the transfer of lipids to a DES is not just a simple substitution of the subphase. In this specific case we have proposed an explanation based on unfolding of the PG head that is enabled by electrostatic screening of the head charges by the charged solvent.

@@

\setlength{\bibsep}{1pt}
\bibliography{bibi}% Produces the bibliography via BibTeX.
\bibliographystyle{rsc}
\end{document}
